\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-05-17 k-planar.tex
Lemma 27 update

\begin{lem}
\label{MewMapGraphLemma}
For every integer $d\geq 3$, every $d$-map graph is a spanning subgraph of a $d$-framed multigraph.
\end{lem}

\begin{proof}
Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a face, and where each vertex of $G_0$ is incident with at most $d$ nations and let $G$ be the corresponding map graph.  Let $G_0^*$ be the dual graph of $G_0$. So the vertices of $G_0^*$ correspond to faces of $G_0$, and each vertex of $G_0^*$ is a nation vertex or a lake vertex.

@@

Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding the nation cycle of each face $F$ of $G_0$, and then triangulating each face with more than $d$ vertices on its boundary.  Each nation cycle of length at least $3$ is now a facial cycle of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary. Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.

By definition, $V(G)=V(\widehat{G}_1)$. To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.  Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary. The vertex $x$ corresponds to a face $F_x$ in $G_0^*$ and the facial cycle of $F_x$ contains $v$ and $w$.  Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$. If $C_x$ has length $2$ then $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.  If $C_x$ has length at least $3$ then $C_x$ bounds a face in $G_1$, so $vw\in E(\widehat{G}_1)$.
\end{proof}

>>>>

\begin{lem}
\label{MewMapGraphLemma}
For every integer $d\geq 3$, every $d$-map graph is a spanning \note{DW}{delete ``spanning''?} 
subgraph of a $d$-framed multigraph.
\end{lem}

\begin{proof}	
Let $G_0$ be a graph embedded in the plane, with each face labelled a nation or a lake, and where each vertex of $G_0$ is incident with at most $d$ nations and let $G$ be the corresponding map graph.

We may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph. Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying $G_0$ is disconnected). Let $v$ and $w$ be vertices in distinct components of the boundary of $f$. Add the edge $vw$ to $G_0$ across $f$. The corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$. Thus no face of $G_0$ is disconnected. Suppose that some face $f$ of $G_0$ has a repeated vertex in the boundary walk of $f$. Let $u,v,w$ be consecutive vertices on the boundary of $f$. So $u,v,w$ are distinct and $vw$ is not an edge in the boundary of $f$. Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake. Since $v$ appears elsewhere in the boundary of $f$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$. Thus no facial walk of $G_0$ has a repeated vertex. Since each facial walk is connected, every face of $G_0$ is bounded by a cycle, and $G_0$ is biconnected.

Let $G_0^*$ be the dual graph of $G_0$. So the vertices of $G_0^*$ correspond to faces of $G_0$, and each vertex of $G_0^*$ is a nation vertex or a lake vertex. Since $G_0$ is biconnected, so is $G^*_0$.

@@

Let $G_1$ be a plane supergraph of $G_0^*$ obtained by adding the nation cycle of each face $F$ of $G_0$, and then triangulating each face with more than $d$ vertices on its boundary. $G_1$ is biconnected since $G^*_0$ is a biconnected spanning subgraph of $G_1$. Each nation cycle of length at least $3$ is now a facial cycle of $G_1$, and no face of $G_1$ has more than $d$ vertices on its boundary. Let $\widehat{G}_1$ be the $d$-framed graph whose frame is $G_1$.

By definition, $V(G)=V(\widehat{G}_1)$ \note{DW}{$V(\widehat{G}_1)=V(G_1)=V(G^*_0)$ which includes vertices corresponding to lakes of $G_0$. So I think this claim is wrong (or am I missing something?). I suggest we write $V(G) \subseteq V(\widehat{G}_1)$ and drop `spanning' from the lemma}. To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.  Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary. The vertex $x$ corresponds to a face $F_x$ in $G_0^*$ and the facial cycle of $F_x$ contains $v$ and $w$.  Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$. If $C_x$ has length $2$ then $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.  If $C_x$ has length at least $3$ then $C_x$ bounds a face in $G_1$, so $vw\in E(\widehat{G}_1)$.
\end{proof}
